Reproducible Data - Assignment One
===================================

###Load required packages
```{r}
	library(knitr)
  library(plyr)
```

###Set the default of echo throughout the document
```{r}
	opts_chunk$set(echo = TRUE)
```
###Read and process the data
```{r}
	loadData <- function(dataURL="", destF="default.csv", method = NULL){
  		if(!file.exists(destF)){
           		temp <- tempfile()
            	download.file(dataURL, temp, method = method)
            	unzip(temp, destF)
            	unlink(temp)
        	}else{
            	message("Data already downloaded.")
        	}
		}

	dataURL <-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
	loadData(dataURL, "activity.csv")


	data <- read.csv("activity.csv", header = TRUE, sep = ',', colClasses = c("numeric", "character",
                                                                          "integer"))
	data$day <- weekdays(as.Date(data$date))
	data$DateTime<- as.POSIXct(data$date, format="%Y-%m-%d")
```
###Check the data
```{r}
	str(data)
	head(data)
```

#Histogram of the total number of steps taken each day

###Total steps per day
```{r}
	sumTable <- aggregate(data$steps ~ data$date, FUN=sum, )
	colnames(sumTable)<- c("Date", "Steps")
```
###Histogram showing total steps per day
```{r}
	hist(sumTable$Steps, breaks=5, xlab="Steps", main = "Steps per day", col="yellow")
```

#Mean and median number of steps taken each day

###Show the mean of steps
```{r}	
	as.integer(mean(sumTable$Steps))
```
###Show the median of steps
```{r}
	as.integer(median(sumTable$Steps))
```

#Time series plot of the average number of steps taken
###Load required packages
```{r}
library(plyr)
library(ggplot2)
```
###Recall data and remove NAs
```{r}	
	clean <- data[!is.na(data$steps),]
```
###Generate average number of steps per interval
```{r}
	intervalTable <- ddply(clean, .(interval), summarize, Avg = mean(steps))
``` 
###Create a line plot showing the average number of steps per interval
```{r}
	p <- ggplot(intervalTable, aes(x=interval, y=Avg), xlab = "Interval", ylab="Average Number of Steps")
	p + geom_line()+xlab("Interval")+ylab("Average Number of Steps")+ggtitle("Average Number of Steps per interval")
```

#The 5-minute interval that, on average, contains the maximum number of steps

###Find which interval has the highest average number of steps
```{r}
	maxSteps <- max(intervalTable$Avg)
	intervalTable[intervalTable$Avg==maxSteps,1]
```

#Describe and show a strategy for imputing missing data

###Find the total number of NAs in the original data
```{r}
	nrow(data[is.na(data$steps),])
```
###Replace NAs using the average 5-minute interval
```{r}
	avgTable <- ddply(clean, .(interval, day), summarize, Avg = mean(steps))
	nadata<- data[is.na(data$steps),]
	newdata<-merge(nadata, avgTable, by=c("interval", "day"))
```
###Reorder the new substituted data in the same format as the clean data
```{r}
	newdata2<- newdata[,c(6,4,1,2,5)]
	colnames(newdata2)<- c("steps", "date", "interval", "day", "DateTime")
```
###Merge the NA averages and non NA data together
```{r}	
	mergeData <- rbind(clean, newdata2)
```
###Create sum of steps per date for comparison
```{r}	
	sumTable2 <- aggregate(mergeData$steps ~ mergeData$date, FUN=sum, )
	colnames(sumTable2)<- c("Date", "Steps")
```
###Mean of Steps without NAs
```{r}	
	as.integer(mean(sumTable2$Steps))
```
###Median of Steps without NAs
```{r}
	as.integer(median(sumTable2$Steps))
```
#Histogram of the total number of steps taken each day after missing values are imputed
```{r}	
	hist(sumTable2$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day", col="green")
	hist(sumTable$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day", col="purple", add=T)
	legend("topright", c("Imputed", "NA"), fill=c("green", "purple") )
```

#Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

###Load the required package
```{r}
library(lattice) 
```
###Create new category based on the days of the week
```{r}
	mergeData$DayCategory <- ifelse(mergeData$day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
```

###Summarize data by interval and type of day
```{r}
	intervalTable2 <- ddply(mergeData, .(interval, DayCategory), summarize, Avg = mean(steps))
```
###Plot data in a panel plot
```{r}	
	xyplot(Avg~interval|DayCategory, data=intervalTable2, type="l",  layout = c(1,2),
	       main="Average steps per 5-minute interval across weekdays and weekends",
	       ylab="Average Number of Steps", xlab="Interval")
```
